---
title: "ESM204 Homework 4"
author: "Nick McManus and Katheryn Moya"
date: "5/21/2022"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(equatiomatic)
```

```{r}
# read in data
warming <- read_csv(here("data", "warming.csv")) %>% 
  select(-1)
damages <- read_csv(here("data", "damages.csv")) #in degrees C
```

### Q1: Estimated Damages

Using `damages.csv`, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature.

-   Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r}
# create new column with squared values to derive quadratic model
damages <- damages %>% 
  mutate(warming_sq = warming^2)

# run regression using the squared and original warming values
damages_model <- lm(damages ~ 0 + warming + warming_sq, data = damages)

# summary(damages_model)

# make quadratic function from the model
damages_fxn <- function(x){
  damages_model$coefficients[1]*x+damages_model$coefficients[2]*x^2
}

```


```{r}
# plot estimated damage function overlaid with scatterplot
ggplot(data = damages, aes(x = warming, y = damages)) +
  geom_point(alpha = 0.3, size = 1.4, color = "sienna")+
  stat_function(fun = damages_fxn, color = "slateblue4", size = 1) +
  theme_classic() +
  labs(x = "\nWarming (degrees C)",
       y = "Damages ($)\n",
       title = "Estimated Damages with Global Temperature Change")
```

### Q2: Damages under different scenarios

Use `warming.csv` and your estimated damage function to predict damages in each year under the Baseline climate and the pulse scenario. Make four plots:

1.  Damages over time without the pulse,

2.  Damages over time with the pulse

3.  The difference in damages over time that arises from the pulse, and

4.  The difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

```{r}
# assign value to one-time pulse of 35 billion tons CO2
pulse <- 35000000000

# create new dataframe with values for damages under baseline and pulse scenarios
warming_damages <- warming %>%
  mutate(damages_baseline = damages_fxn(warming_baseline)) %>%
  mutate(damages_pulse = damages_fxn(warming_pulse)) %>%
  mutate(damages_diff = damages_pulse - damages_baseline) %>%
  mutate(damages_diff_co2 = damages_diff/pulse)


# plot damages over time without pulse
ggplot(data = warming_damages, aes(x = year, y = damages_baseline)) +
  geom_line(color = "slateblue4", size = 1) +
  theme_classic() +
  labs(x = "\nYear",
       y = "Damages ($)\n",
       title = "Estimated Damages without CO2 Pulse") +
    theme(title = element_text(face = "bold"))


# plot damages over time with pulse
ggplot(data = warming_damages, aes(x = year, y = damages_pulse)) +
  geom_line(color = "red", size = 1) +
  theme_classic() +
  labs(x = "\nYear",
       y = "Damages ($)\n",
       title = "Estimated Damages with CO2 Pulse") +
  theme(title = element_text(face = "bold"))


# plot difference in damages over time from pulse
ggplot(data = warming_damages, aes(x = year, y = damages_diff)) +
  geom_line(color = "forestgreen", size = 1) +
  theme_classic() +
  labs(x = "\nYear",
       y = "Difference in damages ($)\n",
       title = "Damages Difference from CO2 Pulse") +
  theme(title = element_text(face = "bold"))


# plot difference in damages over time from pulse per ton CO2
ggplot(data = warming_damages, aes(x = year, y = damages_diff_co2)) +
  geom_line(color = "goldenrod", size = 1) +
  theme_classic() +
  labs(x = "\nYear",
       y = "Difference in damages per ton CO2 ($)\n",
       title = "Damages Differences from Pulse per ton CO2") +
    theme(title = element_text(face = "bold"))
```

### Q3: Social Cost of Carbon

The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

```{r}
# assign value to starting year of 2022
baseline_year <- 2022

# create new dataframe with present values from discount rates from 0 to 5%
warming_damages_scc <- warming_damages %>%
  select(year, damages_diff_co2) %>%
  mutate(pv_0 = damages_diff_co2*((1+0.0)^(-(year-baseline_year)))) %>%
  mutate(pv_1 = damages_diff_co2*((1+0.01)^(-(year-baseline_year)))) %>%
  mutate(pv_2 = damages_diff_co2*((1+0.02)^(-(year-baseline_year)))) %>%
  mutate(pv_3 = damages_diff_co2*((1+0.03)^(-(year-baseline_year)))) %>%
  mutate(pv_4 = damages_diff_co2*((1+0.04)^(-(year-baseline_year)))) %>%
  mutate(pv_5 = damages_diff_co2*((1+0.05)^(-(year-baseline_year)))) %>%
  mutate(pv_6 = damages_diff_co2*((1+0.06)^(-(year-baseline_year)))) %>%
  mutate(pv_7 = damages_diff_co2*((1+0.07)^(-(year-baseline_year)))) %>%
  mutate(pv_8 = damages_diff_co2*((1+0.08)^(-(year-baseline_year)))) %>%
  mutate(pv_9 = damages_diff_co2*((1+0.09)^(-(year-baseline_year)))) %>%
  mutate(pv_10 = damages_diff_co2*((1+0.1)^(-(year-baseline_year))))

discount_rates <- c(0,1,2,3,4,5,6,7,8,9,10)
scc_calc <- c(sum(warming_damages_scc$pv_0),sum(warming_damages_scc$pv_1),
                       sum(warming_damages_scc$pv_2),sum(warming_damages_scc$pv_3),
                       sum(warming_damages_scc$pv_4),sum(warming_damages_scc$pv_5),
                       sum(warming_damages_scc$pv_6),sum(warming_damages_scc$pv_7),
                       sum(warming_damages_scc$pv_8),sum(warming_damages_scc$pv_9),
                       sum(warming_damages_scc$pv_10))

scc_df <- data.frame(discount_rates,
                     scc_calc)

```

```{r}
# plot the discount rates and SCC
ggplot(data = scc_df, aes(x = discount_rates, y = scc_calc)) +
  geom_point(color = "black") +
  geom_smooth(se = 0)+
  theme_classic() +
  labs(x = "\nDiscount rate (%)",
       y = "SCC ($/ton of CO2)\n",
       title = "SCC at Various Discount Rates") +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

Discount rates determine the social cost of carbon. The greater the discount rate, the lower the social cost of carbon is considered to be. Highly discounting the benefits of emission reductions has great implications for climate change mitigation as it decreases motivation for climate policy. This is because a higher discount rate indicates a higher value placed on immediate benefits compared to future benefits, so society is less willing to take on the costs of climate action. In essence, higher discount rates treat the future as less valuable.


### Q4: Ramsey Rule

The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation:$r = ρ + ηg$
Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

```{r}
# find the discount rate (as percent) according to NASEM
r_ramsey <- (0.001 + (2*0.01))*100

# use ramsey discount rate to calculate SCC
warming_damages_ramsey <- warming_damages %>%
  select(year, damages_diff_co2) %>%
  mutate(pv_ramsey = damages_diff_co2*((1+(r_ramsey/100))^(-(year-baseline_year))))
scc_ramsey <- sum(warming_damages_ramsey$pv_ramsey)

```

```{r}
# plot with Ramsey Rule discount rate

ggplot(data = scc_df, aes(x = discount_rates, y = scc_calc)) +
  geom_point(color = "black", size = 2, alpha = 0.3) +
  theme_classic() +
  labs(x = "\nDiscount rate (%)",
       y = "SCC ($/ton of CO2)\n",
       title = "SCC of Ramsey Rule") +
  geom_vline(xintercept = 2.1, color = "slateblue1", linetype = "dashed") +
  geom_point(aes(x = 2.1, y = 71.35), color = "slateblue4", size = 2.5, alpha = 0.8) +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

Using the Ramsey Rule as advised by the National Academies of Sciences, Engineering, and Medicine, the SCC is #`r round(scc_ramsey,2)` with a discount rate of `r r_ramsey`%, as is indicated by the blue point on the graph above.


### Q5: Choosing between two policies

**What is the expected present value of damages up to 2100 under Policy A? What is the expected present value of damages up to 2100 under Policy B?** 
```{r}
# set discount rate for both policies
policy_discount_rate <- 0.02

# Policy A
warming_damages_policy_a <- warming_damages %>%
  select(year, warming_baseline, damages_baseline) %>%
  mutate(damages_2 = damages_fxn(1.5*warming_baseline)) %>%
  mutate(pv_baseline = damages_baseline*(1 + policy_discount_rate)^(-(year - baseline_year))) %>%
  mutate(pv_2 = damages_2*(1 + policy_discount_rate)^(-(year - baseline_year)))

# probability of two scenarios are both 50%
prob_a1 <- 0.5
prob_a2 <- 0.5

# expected present value for both scenarios
exp_pv_a1 <- sum(warming_damages_policy_a$pv_baseline)
exp_pv_a2 <- sum(warming_damages_policy_a$pv_2)

# expected present value of damages up to 2100 
exp_pv_policy_a <- prob_a1 * exp_pv_a1 + prob_a2 * exp_pv_a2
# convert to trillions because number so large
exp_pv_policy_a_trillions <- exp_pv_policy_a/1000000000000

```


```{r}
# Policy B
warming_damages_policy_b <- warming_damages %>%
  select(year, warming_baseline) 
for (i in 30:79){
  warming_damages_policy_b$warming_baseline[i]=1.29
}

warming_damages_policy_b <- warming_damages_policy_b %>%
  mutate(damages_b = damages_fxn(warming_baseline)) %>%
  mutate(pv_b = damages_b*(1 + policy_discount_rate)^(-(year - baseline_year)))

# 100% probability of one scenario under policy B
prob_b <- 1.0

# expected present value of damages up to 2100
exp_pv_policy_b <- sum(warming_damages_policy_b$pv_b)*prob_b
# convert to trillions because number so large
exp_pv_policy_b_trillions <- exp_pv_policy_b/1000000000000
```

Under Policy A (business as usual), the expected present value of damages up to 2100 is $`r round(exp_pv_policy_a_trillions,2)` trillion. Under Policy B (strong, immediate action on climate change), the expected present value of damages up to 2100 is $`r round(exp_pv_policy_b_trillions,2)` trillion.

**Suppose undertaking Policy A costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.**

```{r}
# create fxns for cost policies A and B
policy_a_cost <- function(x){exp_pv_policy_a}
policy_b_cost <- function(x){exp_pv_policy_b + x}

# create matrices
costs <- c(0,1e15,2e15,3e15,4e15,5e15,6e15,7e15,8e15,9e15,10e15)
total_cost_policy_a <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)
total_cost_policy_b <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)

# combine matrices into one dataframe
cost_df <- data.frame(costs, total_cost_policy_a, total_cost_policy_b)
cost_df <- cost_df %>%
  mutate(total_cost_policy_a = policy_a_cost(costs)) %>%
  mutate(total_cost_policy_b = policy_b_cost(costs))

# put dataframe for policy costs in tidy format for ease of plotting
costs_df_long <- cost_df %>%
  pivot_longer(cols = c(total_cost_policy_a,total_cost_policy_b), names_to = 'policy', values_to = 'total_cost') %>%
  mutate(policy = case_when(
    policy == "total_cost_policy_a" ~ "policy a",
    policy == "total_cost_policy_b" ~ "policy b"))

ggplot(data = costs_df_long, aes(x = costs, y = total_cost)) +
  geom_line(aes(color = policy), size = 1) +
  theme_classic() +
  labs(x = "Initial Cost of Policy B ($)",
       y = "Expected Value of Policy ($)",
       title = "Expected Value vs. Initial Cost of Policy") +
  theme(title = element_text(face = "bold"))


#find point where policy a cost = policy b cost using uniroot()
dif_costs_fun <- function(x){exp_pv_policy_a - (exp_pv_policy_b + x)}
equal_cost <- uniroot(dif_costs_fun, c(1e15,2.5e15))$root
```

Using the initial cost of Policy B as $`r equal_cost` both policies will have equivalent expected values. If Policy B were to cost any less than that initial cost, Congress should undertake Policy B. On the other hand, if Policy B were to cost more than this initial amount, Congress should undertake Policy A.


### Q6: Risk Aversion

**Now suppose that society is risk averse. How do you think this would change your advice to Congress?**

A risk averse society would prefer Policy B because it projects lower future damages as opposed to the high cost outcome of Policy A. It would be necessary to tell Congress to evaluate expected damages of each policy by plugging in each policy’s estimated costs to the function used to determine society’s utility. In order to determine the true cost of each policy, Congress should add the costs of each individual policy to their respective expected damages. This will allow Congress to determine which policy has the lowest expected costs, which is preferred by a risk averse society.

